The 2014 SMT Competition

### 2015-07-30 2014/report/SMTCOMP.tex
Edits.
Significant reorganization of the sections

\davidc{Check that the outline is still correct}

\section{Context of the Competition: SMT-LIB language and benchmarks}
\label{sec:context}

@@

As stated, the goal of SMT-COMP is similar to the goal of the SAT competition. The SMT benchmarks extend SAT by incorporating defined theories, such as the theory of arrays, or of undefined functions, or of arithmetic, or of bit-vectors. In addition, there may be constraints on the set of expressions allowed, such as only linear arithmetic, or only integer difference arithmetic. The theories also define \emph{sorts}, which make the theories a kind of typed first-order logic. Examples of sorts used in current theories are Boolean, Int, Real, bit-vectors of various lengths, and arrays with arbitrary sorts as index and value. Each combination of underlying theories and language constraints is a \emph{logic}.
The names of the logics as used in SMT-LIB are cryptic combinations of initials. For example, AUFLIA is the logic with a
combination of Arrays (A), Uninterpreted Functions (UF), and linear integer arithmetic (LIA). Table \ref{logicAbbreviations} can be used to interpret these names.
\begin{table}
\begin{tabular}{|l|l|}
\hline
QF\_ & Quantifier-Free \\
A & Arrays \\
UF & Unidentified Functions \\
BV & Bit-Vector \\
L & Linear (arithmetic) \\
N & Non-Linear (arithmetic) \\
IA/RA/IRA & Integer/Real/Both arithmetic \\
DL & Difference Logic \\
\hline
\end{tabular}
\caption{Abbreviations used in logic names}
\label{logicAbbreviations}
\end{table}

@@

The language includes commands that allow a fine-grained interaction
with the solver, whereby client tools may incrementally push and pop
symbol declarations and assertions while running various
satisfiability checks, inspecting models, or obtaining unsatisfiability
proofs. Benchmarks problems that have more than one
\lstinline{check-sat} command are called \emph{application} or \emph{incremental
  benchmarks}. Not all SMT solvers support all of these interaction
facilities, and the main track of the competition does not use such
application benchmarks.

As of June 2014, the SMT-LIB repository contained over 130,000
non-incremental benchmarks divided into 34 background theories, and
close to 10,000 incremental benchmarks distributed across 8 background
theories. The size of a benchmark file may vary from a few hundred
bytes to several gigabytes. Fig.~\ref{fig:smtlib-bysize} shows the
distribution of the benchmarks by file size.

@@

In planning the 2014 competition, the organizers' overall goal was to encourage breadth
in the capability of SMT solvers. SMT-COMP 2014 benefited from the evaluation that was performed in 2013 and the experience of previous competitions. As a result we established the following emphases:
\begin{itemize}
\item In 2012 we narrowed the competition to a smaller number of more significant logics. In response to feedback, in 2014 we reverted to the practice of evaluating solvers in all available divisions.
\item We ran the competition with all available benchmarks. To be eligible, a benchmark must be well-formed and have a known answer. We were able to use all benchmarks because we had nearly exclusive use of the StarExec cluster during the time we were executing SMT-COMP. This point is discussed further in \S\ref{sec:procedure}.
\item In 2012 some experimental tracks were held: parallel performance, unsat cores, and proof generation. The participation in those tracks was light. Since in 2014 we also had to migrate to using StarExec, we held only a main track and an application (incremental) track in the 2014 competition. The organizers still appreciate the value of measuring the performance of solvers on new features such as model generation, unsat core determination, and proof generation, and recommend that they be reinstated in some future edition of the competition.
\item A significant change in 2014 was to increase the time-out limit for a solver processing a given benchmark from 25 minutes (in 2012) to 40 minutes.
\item A significant result of the 2013 Evaluation was that the result of a competition was highly sensitive to the selection of benchmarks: different random selections of benchmarks resulted in different winning orders in a large fraction of samples. Hence, an aim for 2014 was to use as large a benchmark set as possible in the competition to minimize this effect.

@@

@@

@@

StarExec consists of 192 2.4\,GHz computational nodes, running Red Hat Linux~6.3. Each node has a dual-processor, quad-core CPU (model Intel Xeon CPU E5-2609 at 2.4GHz), 10240 KB cache, and either 129022\,MB (32 nodes) or 258294\,MB of memory (160 nodes). In competition, processes were limited to 100\,GB memory. To avoid interference among job-pairs, only one job-pair was executed at a time on each node.

For comparison, the previous computational cluster (SMT-Exec) had just 11 nodes:
\begin{itemize}
\item Nine are 2.4\,GHz AMD Opteron 250s, configured for single core, 64-bit processing, 1 MB cache, 4 GB main memory
\item Two are 2.53\,GHz Intel Xeon E5540s, configured for single core, 64-bit processing, 8 MB cache, 12 GB main memory
\end{itemize}
All eleven execution nodes have the same software configuration:
\begin{itemize}
\item Linux, kernel version 2.6.18-128.2.1.el5
\item Red Hat Enterprise Linux Client release 5.2 (Tikanga)
\item glibc-2.5-24.el5\_2.2
\item runtime libraries from gcc-4.1.2-42.el5
\end{itemize}
The raw CPU speed is not significantly different, but memory size, number of cores, and number of nodes is greatly increased.

@@

@@

An important issue is how solver errors should be taken into account into the metrics. A solver error indicates an unsoundness in the tool, which is highly undesirable; however, highly penalizing errors may discourage new entrants whose tools are not yet mature. Different competitions vary on this aspect. For the SAT competition, an error disqualifies a tool in the category it occurs\footnote{See \url{http://satcompetition.org/2014/rules.shtml}.}. In contrast, in the SV-competition~\cite{svcomp-2015}, successful runs account positively towards the final score, while errors count negatively (with an increased weight).

Note that incorrect results are very rare in SMT-COMP, but do occur; for almost all solvers and divisions the value of $e_i$ is 0 and the competition hinges on the values of $c_i$. The speed of the solver is important in two ways. First, if the solver is slow, it will time out before a solution is found and thus the value of $c_i$ will be lower. Second, if there is a tie in the number of errors and correctly solved problems, the total time taken on the correctly solved problems is used as the tie-breaker (even if the solvers solve different subsets of benchmarks); this is a rare occurrence but does happen if, for example, all the benchmark problems in a division are solvable within the time limit---the only situation in the competition in which tie-breaking has been needed. Other competitions (e.g., SAT and SV-COMP) also award points for successful solutions and use the time to solve problems only as a tie-breaker, if at all.

@@

This year's competition also saw the inauguration of a new, sibling competition. The organizers helped start and execute SL-COMP~\cite{SLCOMP}, a competition among solvers for separation logic.

>>>>

\davidc{Correct the outline}

In planning the 2014 competition, the organizers' overall goal was to encourage breadth
in the capability of SMT solvers. SMT-COMP 2014 benefited from the evaluation that was performed in 2013 and the experience of previous competitions. As a result we established the following emphases:
\begin{itemize}
\item In 2012 the competition was narrowed to a smaller number of more significant logics. In response to feedback, in 2014 we reverted to the practice of evaluating solvers in all available divisions.
\item We ran the competition with all available, eligible benchmarks. To be eligible, a benchmark must be well-formed and have a known answer. We were able to use all benchmarks because we had nearly exclusive use of the StarExec computational cluster during the time we were executing SMT-COMP. This point is discussed further in \S\ref{sec:procedure}. \
davidc{Correct reference? maybe to the benchmark section?}
\item In 2012 some experimental tracks were held: parallel performance, unsat cores, and proof generation. The participation in those tracks was light. Since in 2014 we also had to migrate to using the new StarExec cluster to execute the competition, we held only a main track and an application (incremental) track in the 2014 competition. The organizers still appreciate the value of measuring the performance of solvers on new features such as model generation, unsat core determination, and proof generation, and recommend that they be reinstated in some future edition of the competition.
\item A significant change in 2014 was to increase the time-out limit for a solver processing a given benchmark from 25 minutes (in 2012) to 40 minutes. (cf. section \ref{TBD}) \davidc{Add section reference}
\item A significant result of the 2013 Evaluation was that the result of a competition was highly sensitive to the selection of benchmarks: different random selections of benchmarks resulted in different winning orders in a large fraction of samples. Hence, an aim for 2014 was to use as large a benchmark set as possible in the competition to minimize this effect.

\section{SMT-LIB logic, language and solvers}
\label{sec:context}

@@

As stated, the goal of SMT-COMP is similar to the goal of the SAT competition. The SMT logic extends SAT by incorporating defined theories, such as the theory of arrays, or of undefined functions, or of arithmetic, or of bit-vectors. In addition, there may be constraints on the set of expressions allowed, such as only linear arithmetic, or only integer difference arithmetic. The theories also define \emph{sorts}, which make the theories a kind of typed first-order logic. Examples of sorts used in current theories are Boolean, Int, Real, bit-vectors of various lengths, and arrays with arbitrary sorts as index and value. Each combination of underlying theories and language constraints is a \emph{logic}.
The names of the logics as used in SMT-LIB are cryptic combinations of initials. For example, AUFLIA is the logic with a
combination of Arrays (A), Uninterpreted Functions (UF), and linear integer arithmetic (LIA). Table \ref{logicAbbreviations} can be used to interpret these names.
\begin{table}
\begin{tabular}{|l|l|}
\hline
QF\_ & Quantifier-Free \\
A & Arrays \\
UF & Unidentified Functions \\
BV & Bit-Vector \\
L & Linear (arithmetic) \\
N & Non-Linear (arithmetic) \\
IA/RA/IRA & Integer/Real/Both arithmetic \\
DL & Difference Logic \\
\hline
\end{tabular}
\caption{Abbreviations used in logic names}
\label{logicAbbreviations}
\end{table}

@@

@@

@@

As of June 2014, the SMT-LIB repository contained over 130,000
main-track benchmarks divided into 34 background theories, and
close to 10,000 incremental benchmarks distributed across 8 background
theories. The size of a benchmark file may vary from a few hundred
bytes to several gigabytes. Fig.~\ref{fig:smtlib-bysize} shows the
distribution of the benchmarks by file size.

@@

\paragraph{Application benchmarks.} The language includes commands that allow a fine-grained interaction
with the solver, whereby client tools may incrementally push and pop
symbol declarations and assertions while running various
satisfiability checks, inspecting models, or obtaining unsatisfiability
proofs. Benchmarks problems that have more than one
\lstinline{check-sat} command are called \emph{application} or \emph{incremental
  benchmarks}. Not all SMT solvers support all of these interaction
facilities, and the main track of the competition does not use such
application benchmarks.

@@

StarExec consists of 192 2.4\,GHz computational nodes, running Red Hat Linux~6.3. Each node has a dual-processor, quad-core CPU (model Intel Xeon CPU E5-2609 at 2.4GHz), 10240 KB cache, and either 129022\,MB (32 nodes) or 258294\,MB of memory (160 nodes). In competition, processes were limited to 100\,GB memory. To avoid interference among job-pairs, only one job-pair was executed at a time on each node. The software configuration contains
\begin{itemize}[noitemsep,nolistsep]
\item Linux kernel 2.6.32-573.1.1.el6.x86\_64
\item GNU C Library stable release version 2.12
\item gcc (GCC) 4.4.7 20120313 (Red Hat 4.4.7-11)
\end{itemize}

For comparison, the previous computational cluster (SMT-Exec) had just 11 nodes:
\begin{itemize}[noitemsep,nolistsep]
\item Nine are 2.4\,GHz AMD Opteron 250s, configured for single core, 64-bit processing, 1 MB cache, 4 GB main memory
\item Two are 2.53\,GHz Intel Xeon E5540s, configured for single core, 64-bit processing, 8 MB cache, 12 GB main memory
\end{itemize}
All eleven execution nodes have the same software configuration:
\begin{itemize}[noitemsep,nolistsep]
\item Linux, kernel version 2.6.18-128.2.1.el5
\item Red Hat Enterprise Linux Client release 5.2 (Tikanga)
\item glibc-2.5-24.el5\_2.2
\item runtime libraries from gcc-4.1.2-42.el5
\end{itemize}
The raw CPU speed is not significantly different, but memory size, number of cores, and number of nodes is greatly increased.

@@

\section{Other Competitions}
\label{sec:OtherCompetitions}

Competitions among tools for a specific purpose are now common. Indeed SMT-COMP participated among dozen or so tools in the
FLoC Olympic Games in 2014~\cite{TBD}. The competitions differ, of course, in their target problem set, but also differ in
their overall goals, scoring policies, and organization.

The organizers of SMT-COMP helped organize and execute the first edition of SL-COMP~\cite{SLCOMP} as a sibling competition. SL-COMP is a competition among solvers for separation logic. The formal logical underpinnings of SL-COMP are still being 
defined and the relationship between the logic used in SL-COMP and the SMZT-LIB logic is under discussion. However, SL-COMP
did express its benchmark problems in the same syntax as SMT-COMP, used the same StarExec hardware, and used similar 
organization and scoring rules.

\davidc{Update the reference for SLCOMP}

Besides the subject material, the most interesting difference among competitions is in scoring. (The scoring procedure for
SMT-COMP is described in detail in section \ref{sec:floc}.)
Competitions are currently mostly focused on raw solver capability---that is, the numbers and difficulty of problems that can be solved. Thus competition winners are primarily determined by the total number of problems solved correctly,
with the time taken to produce the solutions used only as a tie-breaker. This is the case for SMT-COMP, SAT, and SV-COMP,
for example.

A more significant issue is how solver errors are taken into account in the metrics. A solver error indicates an unsoundness in the tool, which is highly undesirable; however, highly penalizing errors may discourage new entrants whose tools are not yet mature. Different competitions vary on this aspect. SMT-COMP has historically emphasized soundness; a
single error in a division caused the solver to be scored lower than any other solver with no errors, even if the other solvers solved very few problems. For the SAT competition, an error outright disqualifies a tool in the category it occurs\footnote{See \url{http://satcompetition.org/2014/rules.shtml}.}. In contrast, in the SV-competition~\cite{svcomp-2015}, successful runs account positively towards the final score, while errors count negatively (with an increased weight).

\davidc{Research a wider set of competitions}

@@

Note that incorrect results are very rare in SMT-COMP, but do occur; for almost all solvers and divisions the value of $e_i$ is 0 and the competition hinges on the values of $c_i$. The speed of the solver is important in two ways. First, if the solver is slow, it will time out before a solution is found and thus the value of $c_i$ will be lower. Second, if there is a tie in the number of errors and correctly solved problems, the total time taken on the correctly solved problems is used as the tie-breaker (even if the solvers solve different subsets of benchmarks); this is a rare occurrence but does happen if, for example, all the benchmark problems in a division are solvable within the time limit---the only situation in the competition in which tie-breaking has been needed.

@@
